# Results

```{r, warning = FALSE, results = 'hide', message = FALSE}
# install.packages("datasets")
# install.packages("ggplot2")
# install.packages("dplyr")
# install.packages("mi")
# install.packages("naniar")
# install.packages("vcdExtra")
#install.packages("ggalluvial")
#install.packages("maps")
#install.packages("mapproj")
library(datasets) 
library(ggplot2) 
library(dplyr) 
library(data.table)
library(tidyverse)
library(mi)
library(magrittr)
library(plyr)
library(naniar)
library(vcdExtra)
library(ggalluvial)
library (maps)
library (mapproj)
```

## Migration of Individuals

Migration flow within the United States is driven by several competing forces. Although it is all happening in the same country and it varies from State to State, it also affects Age groups and Income levels differently. 

During the 2016-2020 period, we observe the majority of individuals staying within their county.

Based on the chart below, the **most transient group is composed of individuals making between `$1 to $10,000`** as they moved out, moved around, and moved into a different state. Our assumptions are that this could be due to having bigger needs to look for better employment opportunities and they might also have the flexibility to move around more if they do not have stable/long term contracts. On the other hand, **the most stationary group is individuals making `$200,000 or more` per year**. This can be tied to the fact that they might not need to change their primary residence, are comfortable where they live and if desired, they could have the discretionary income to spend time at their second home (i.e. during COVID-19). 


```{r, echo=FALSE, warning=FALSE}
df <- read.csv("migration_data/df_preprocessed.csv")

#Convert to groups/unordered factors
factors_unordered <- c("Status", "State", "State_name")

for (i in factors_unordered){
    df[, i] <- as.factor(df[, i])
}

##Convert age groups to an ordered factor
factors_ordered <- unique(df[,"Age"])

df[,"Age"] <- factor(df[,"Age"], ordered=TRUE, levels=c("under 26", "26 under 35", "35 under 45","45 under 55","55 under 65", "65 and over"))

df[,"GTI"] <- factor(df[,"GTI"], ordered=TRUE, levels=c("$1 under $10,000", "$10,000 under $25,000", "$25,000 under $50,000", "$50,000 under $75,000", "$75,000 under $100,000", "$100,000 under $200,000", "$200,000 or more"))

```

```{r, echo=FALSE, warning=FALSE}
percentages <- function(df){
  df$ind_perc <- (df$Individuals/sum(df$Individuals))
  return(df)
}
```


```{r, echo=FALSE, warning=FALSE}
#Migration flow 2015-2020 per GTI
df_sum <- ddply(df[,c("GTI", "Status", "Individuals")], .(GTI), percentages)

df_sum <- as.data.table(df_sum)[, sum(ind_perc), by = .(GTI, Status)]


ggplot(df_sum, aes(fill=fct_rev(Status), x=GTI,
                         y=V1)) +
  geom_col(position = "fill", alpha = .75) +
  ylab("Migration Status") +
  ggtitle('Migration Status per Gross Total Income 2015-2020') +
  xlab('Gross Total Income') +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme_classic() +
  theme(legend.position = "top", legend.title = element_blank())
```


Comparing the type of migration across age groups reveals insights that are more in line with our previous assumptions. Younger people tend to migrate more than older adults. There seems to be a negative relationship between age and moving. 



```{r, echo=FALSE, warning=FALSE}
#Migration flow 2015-2020 per GTI
df_sum <- ddply(df[,c("Individuals", "Status","Age")], .(Age), percentages)

df_sum <- as.data.table(df_sum)[, sum(ind_perc), by = .(Age, Status)]

ggplot(df_sum, aes(fill=Status, x=Age,
                         y=V1)) +
  geom_col(position = "fill", alpha = .75) +
  ylab('Age Group') +
  ggtitle('Migration Status by Age 2015-2020') +
  xlab("Migration Status") +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme_classic() +
  theme(legend.position = "top", legend.title = element_blank())
```


## Migration and Average Gross Income

Since the data set is quite large, we have considered a subset for the analysis of the Average gross income of the states for the migrants who have migrated out of the state over all the age groups:

For the age group under 26, we can observe that the Average gross income of the migrants has increased in the states of California, New York, and Texas with Texas recording the biggest increase in 2020. This is explored by looking at the concentration within states in a geographic heat map. 

```{r}
#We will now read the file
mig_data <- read.csv("df.csv")


#import the state data
states <- map_data("state")

#convert the data to lower case
mig_data$State_name <- tolower(mig_data$State_name)
states$region <- tolower(states$region)

#Change column name
colnames(states)[5] <- "State_name"
#Merging
states_merge <- merge(mig_data , states , sort = FALSE , by = "State_name")

#filtering out data
df <- states_merge %>% filter(Status == "Outflow")
df$AGI_Year1 <- as.numeric(df$AGI_Year1) / 1000
df0 <- df %>% filter(Age == "under 26")
df1 <- df %>% filter(Age == "26 under 35")
df2 <- df %>% filter(Age == "35 under 45")
df3 <- df %>% filter(Age == "45 under 55")
df4 <- df %>% filter(Age == "65 and over")
df5 <- df %>% filter(Age == "55 under 65")

g0<-ggplot(df0, aes (long, lat)) + geom_polygon(aes(group=group, fill= AGI_Year1))+ scale_fill_continuous(low = "cadetblue", high = "blue") +coord_map ()
g0 <- g0 + facet_wrap(~factor(Year))
g0 <- g0 + labs(x="Longitude", y="Latitude", title=" Average Gross Income vs  Region (Under 26)") + labs(fill='AGI(USD in millions)')
g0
```

```{r}
g1<-ggplot(df1, aes (long, lat)) + geom_polygon(aes(group=group, fill= AGI_Year1))+ scale_fill_continuous(low = "cadetblue", high = "blue") +coord_map ()
g1 <- g1 + facet_wrap(~factor(Year))
g1 <- g1 + labs(x="Longitude", y="Latitude", title=" Average Gross Income vs  Region (Age 26-35)") + labs(fill='AGI(USD in millions)')
g1
```


For the age group of individuals between 26 and 35, we can observe that the Average gross income of the migrants has increased in the states of California, New York, and Texas from 2016 to 2017 but there has been a decrease in California and Texas from 2017 onwards. The AGI of New York for this category has increased continuously over the years from 2016 to 2020.

The average gross income of the migrants between 35 and 45, has decreased over the years in the states of California with Texas recording similar figures over the years. From the graph, we can see that there has been a sudden increase in the average gross income from 2019 to 2020.

```{r}
g2<-ggplot(df2, aes (long, lat)) + geom_polygon(aes(group=group, fill= AGI_Year1))+ scale_fill_continuous(low = "cadetblue", high = "blue")+coord_map ()
g2 <- g2 + facet_wrap(~factor(Year))
g2 <- g2 + labs(x="Longitude", y="Latitude", title=" Average Gross Income vs  Region (Age 35-45)") + labs(fill='AGI(USD in millions)')
g2
```


```{r}
g3<-ggplot(df3, aes (long, lat)) + geom_polygon(aes(group=group, fill= AGI_Year1))+ scale_fill_continuous(low = "cadetblue", high = "blue") +coord_map ()
g3 <- g3 + facet_wrap(~factor(Year))
g3 <- g3 + labs(x="Longitude", y="Latitude", title=" Average Gross Income vs  Region (Age 45-55)") + labs(fill='AGI(USD in millions)')
g3
```


For the age group of individuals between 45 and 55, we can observe that the Average gross income of the migrants has decreased over the years in the states of California with Texas recording similar figures until 2019 and then dropping in 2020. In this category, we can also observe that Illinois is also a significant contributor over other states. From the graph, we can see that there has been an incremental increase in the average gross income from 2018 to 2020


As individuals get closer to retiring, starting by 55 and 65 age groups, we can observe that Florida has a relatively high average gross income as compared to other states. New York is another state that has a relatively high average gross income. In the 65 years and above range, we can observe that the Average gross income of the migrants has decreased over the years in the states of California with Texas recording similar figures over the years and then dropping in 2020. From the graph, we can see that there has been an incremental increase in the average gross income from 2019 to 2020.

```{r}
g4<-ggplot(df4, aes (long, lat)) + geom_polygon(aes(group=group, fill= AGI_Year1))+ scale_fill_continuous(low = "cadetblue", high = "blue") +coord_map ()
g4 <- g4 + facet_wrap(~factor(Year))
g4 <- g4 + labs(x="Longitude", y="Latitude", title=" Average Gross Income vs  Region (Age 55-65)") + labs(fill='AGI(USD in millions)')
g4 <- g4 + theme(legend.key.size = unit(0.7, 'cm'))
g4
```


```{r}
g5<-ggplot(df3, aes (long, lat)) + geom_polygon(aes(group=group, fill= AGI_Year1))+ scale_fill_continuous(low = "cadetblue", high = "blue") +coord_map ()
g5 <- g5 + facet_wrap(~factor(Year))
g5 <- g5 + labs(x="Longitude", y="Latitude", title=" Average Gross Income vs  Region (Age 65+)") + labs(fill='AGI(USD in millions)')
g5 <- g5 + theme(legend.key.size = unit(0.7, 'cm'))
g5
```




## Number of Tax Returns


```{r, echo=FALSE, warning=FALSE}
mig_data <- read.csv("migration_data/df.csv")
mig_data <- mig_data %>% select(-X)

# Combine 
mig_grouped <- as.data.table(mig_data)[, sum(Returns), by = .(Age, Status, State_name)]

# assigning the column name to a new name
colnames(mig_grouped)[4] <- "Total_returns"
colnames(mig_grouped)[3] <- "State_name"

top_total_states <- c("California", "Texas", "Florida", "New York", "Pennsylvania", "Illinois")

# Filter plotting data
status_vec <- c("Inflow", "Outflow", "Same State")

mig_grouped <- mig_grouped %>% 
  filter(Status %in% status_vec) %>%
  filter(State_name %in% top_total_states)

mig_grouped$Total_returns <- as.numeric(mig_grouped$Total_returns) / 1000

ggplot(mig_grouped, aes(x = reorder(State_name, desc(Total_returns)), y = Total_returns, fill = Status)) +
  geom_bar(position = "dodge", stat = "identity") +
  facet_wrap(~factor(Age, levels=c('under 26', '26 under 35', '35 under 45', '45 under 55', '55 under 65', '65 and over'))) +
  # formatting
  ggtitle("State & age wise inflow and outflow comparsion") +
  labs(y = "Number of Returns in thousands", x = "States") +
  theme(plot.title = element_text(face = "bold")) +
  theme(plot.subtitle = element_text(face = "bold", color = "grey35")) +
  theme(plot.caption = element_text(color = "grey68")) + 
  theme(axis.text = element_text(size = 4))

```


The following observations can be made from the above graph which is a facet graph of the number of returns with respect to the age category (The number of returns is the number of households that migrated):

- We can observe that among all the groups, the same state migration dominates interstate migration. However, in states like Florida, senior individuals especially the ones above the age of 55 prefer to move to another state than move to a different county in the same state

- The majority of the migrants are in the age group of 25 and 35. These individuals might be moving in search of better career and employment opportunities.



```{r, echo=FALSE, warning=FALSE}
mig_data <- read.csv("migration_data/df.csv")
mig_data <- mig_data %>% select(-X)

colnames(mig_data)[9] <- "Gross_total_income"

# Combine
mig_grouped <- as.data.table(mig_data)[, sum(Returns), by = .(Gross_total_income, Status, State_name)]


# assigning the column name to a new name
colnames(mig_grouped)[4] <- "Total_returns"
colnames(mig_grouped)[3] <- "State_name"

top_total_states <- c("California", "Texas", "Florida", "New York", "Pennsylvania", "Illinois")

# Filter plotting data
status_vec <- c("Inflow", "Outflow", "Same State")

mig_grouped <- mig_grouped %>% 
  filter(Status %in% status_vec) %>%
  filter(State_name %in% top_total_states)

mig_grouped$Total_returns <- as.numeric(mig_grouped$Total_returns) / 1000

ggplot(mig_grouped, aes(x = reorder(State_name, desc(Total_returns)), y = Total_returns, fill = Status)) +
  geom_bar(position = "dodge", stat = "identity") +
  facet_wrap(~factor(Gross_total_income,  levels=c('$1 under $10,000', '$10,000 under $25,000', '$25,000 under $50,000', '$50,000 under $75,000', '$75,000 under $100,000', '$100,000 under $200,000', '$200,000 or more'))) +
  # formatting
  ggtitle("State & Gross Income wise inflow and outflow comparsion") +
  labs(y = "Number of Returns in thousands", x = "States") +
  theme(plot.title = element_text(face = "bold")) +
  theme(plot.subtitle = element_text(face = "bold", color = "grey35")) +
  theme(plot.caption = element_text(color = "grey68")) + 
  theme(axis.text = element_text(size = 4))

```

The following observations can be made from the above graph which is a facet graph of the number of returns with respect to income of households ( The number of returns is the number of households that migrated):

- In most of the income categories, the same-state migration dominates the inflow and outflow migration.

- Maximum migration has been observed in the 25000 USD – 50000 USD income category

- Least migration has been observed in the 200,000 USD and above income category

- In states like Texas and Florida compared to all the other states, the inflow migration is greater than the outflow migration which means that more individuals are entering these states than exiting



## Temporal Analysis | Migration by Year


```{r, echo=FALSE, warning=FALSE}
#We will now read the file
mig_data <- read.csv("migration_data/df.csv")
mig_data <- mig_data %>% select(-X)

colnames(mig_data)[9] <- "Gross_total_income"

# Combine
mig_grouped <- as.data.table(mig_data)[, sum(Returns), by = .(Year, Status, State_name)]

# assigning the column name to a new name
colnames(mig_grouped)[4] <- "Total_returns"
colnames(mig_grouped)[3] <- "State_name"

top_total_states <- c("California", "Texas", "Florida", "New York", "Pennsylvania", "Illinois")

# Filter plotting data
status_vec <- c("Inflow", "Outflow", "Same State")

mig_grouped <- mig_grouped %>% 
  filter(Status %in% status_vec) %>%
  filter(State_name %in% top_total_states)

mig_grouped$Total_returns <- as.numeric(mig_grouped$Total_returns) / 1000

ggplot(mig_grouped, aes(x = reorder(State_name, desc(Total_returns)), y = Total_returns, fill = Status)) +
  geom_bar(position = "dodge", stat = "identity") +
  facet_wrap(~(Year)) +
  # formatting
  ggtitle("State & Year wise inflow and outflow comparsion") +
  labs(y = "Number of Returns in thousands", x = "States") +
  theme(plot.title = element_text(face = "bold")) +
  theme(plot.subtitle = element_text(face = "bold", color = "grey35")) +
  theme(plot.caption = element_text(color = "grey68"))+ 
  theme(axis.text = element_text(size = 4))


```


The following observations can be made from the above graph which is a facet graph of the number of returns with respect to the year

- Over the span of 5 years, migration is highest in the states  California and Texas compared to all the other states. One possible reason could be the low tax rates in Texas and the vast variety of opportunities( like software development) available in California.

- Over the span of 5 years, inflow migration is greater than outflow migration in Texas and Florida and it’s the other way in all the other states (Top 5 states with migration figures)

- All 3 types of migration are high in all the top 5 states in 2017 compared to the other years

- In all the states, Same state migration dominates inflow and outflow migration


```{r, echo=FALSE, warning=FALSE}
#We will now read the file
mig_data <- read.csv("migration_data/df.csv")

colnames(mig_data)[9] <- "Gross_total_income"

# Combine
mig_grouped <- as.data.table(mig_data)[, sum(Returns), by = .(Year, Status, State_name)]

# assigning the second column name to a new name
colnames(mig_grouped)[4] <- "Total_returns"
colnames(mig_grouped)[3] <- "State_name"

top_total_states <- c("California", "Texas", "Florida", "New York", "Pennsylvania", "Illinois")

# Filter plotting data
status_vec <- c("Inflow", "Outflow", "Same State")

mig_grouped <- mig_grouped %>% 
  filter(Status %in% status_vec) %>%
  filter(State_name %in% top_total_states) %>%
  mutate(State_name = forcats::fct_reorder2(State_name, Year, Total_returns))

mig_grouped$Total_returns <- as.numeric(mig_grouped$Total_returns) / 1000

ggplot(mig_grouped, aes(Year, Total_returns, color = State_name)) + geom_line() +
  ggtitle("Total Returns from 2016 to 2020") +
  facet_wrap(~factor(Status)) +
  labs (x = "Year", y = "Total Returnsin thousands") +
  theme_grey(16) +
  theme(legend.title = element_blank())+ 
  theme(axis.text = element_text(size = 5))+
  theme(axis.title=element_text(size=10),plot.title=element_text(size=13))

```


The following observations can be made from the above graph which is a facet graph of the number of returns with respect to the type of migration over the span of 5 years

- In all the states, migration peaked in 2017 and then reduced over the next few years and it again increased in 2019. One of the reasons for a possible hike in migration patterns in 2017 could be the introduction of new migration laws in 2017 by the US government.

- Over the span of 5 years from 2016 to 2020, Florida has the maximum inflow migration , California has the maximum outflow migration and California and Texas have the maximum same state migration.


```{r, echo=FALSE, warning=FALSE}
#We will now read the file
mig_data <- read.csv("df.csv")

colnames(mig_data)[9] <- "Gross_total_income"

# Combine
mig_grouped <- as.data.table(mig_data)[, sum(Returns), by = .(Year, Status, State_name)]

# assigning the second column name to a new name
colnames(mig_grouped)[4] <- "Total_returns"
colnames(mig_grouped)[3] <- "State_name"

top_total_states <- c("California", "Texas", "Florida", "New York", "Pennsylvania", "Illinois")

# Filter plotting data
status_vec <- c("Non-migrant" )

mig_grouped <- mig_grouped %>% 
  filter(Status %in% status_vec) %>%
  filter(State_name %in% top_total_states) %>%
  mutate(State_name = forcats::fct_reorder2(State_name, Year, Total_returns))

mig_grouped$Total_returns <- as.numeric(mig_grouped$Total_returns) / 1000

ggplot(mig_grouped, aes(Year, Total_returns, color = State_name)) + geom_line() +
  ggtitle("Total Returns from 2016 to 2020") +
  facet_wrap(~factor(Status)) +
  labs (x = "Year", y = "Total Returns in thousands") +
  theme_grey(16) +
  theme(legend.title = element_blank()) +
  theme(axis.title=element_text(size=10),plot.title=element_text(size=13))


```



The following observations can be made from the above graph which is a  time series line that shows the number of non-migrant households over the span of 5 years.

- There is a dip in the number of people who did not migrate in the year 2017

- This is in accordance with the previous graph that shows an increase in migration in 2017 compared to all other years from 2017 to 2020.



<!-- #We will now read the file -->
<!-- mig_data <- read.csv("df.csv") -->

<!-- # create a theme for dot plots, which can be reused -->
<!-- theme_dotplot <- theme_bw(14) + -->
<!--   theme(axis.text.y = element_text(size = rel(.75)), -->
<!--         axis.ticks.y = element_blank(), -->
<!--         axis.title.x = element_text(size = rel(.75)), -->
<!--         panel.grid.major.x = element_blank(), -->
<!--         panel.grid.major.y = element_line(size = 0.5), -->
<!--         panel.grid.minor.x = element_blank()) -->

<!-- colnames(mig_data)[9] <- "Gross_total_income" -->

<!-- # Combine  -->
<!-- mig_grouped <- as.data.table(mig_data)[, sum(Returns), by = .(Year, Status, State_name)] -->

<!-- # assigning the second column name to a new name -->
<!-- colnames(mig_grouped)[4] <- "Total_returns" -->
<!-- colnames(mig_grouped)[3] <- "State_name" -->


<!-- # Filter plotting data -->
<!-- status_vec <- c("Inflow", "Outflow", "Same State") -->

<!-- mig_grouped <- mig_grouped %>%  -->
<!--   filter(Status %in% status_vec) %>% -->
<!--   filter(State_name != "Total US")  -->

<!-- ggplot(mig_grouped, -->
<!--        aes(Total_returns, fct_reorder2(State_name, Status, Total_returns, .desc = FALSE), -->
<!--            color = Status)) + -->
<!--   geom_point() + ggtitle("total returns for top 10 states facet by year") + ylab("") + -->
<!--   theme_dotplot + -->
<!--   facet_wrap(~factor(Year)) -->



```{r, echo=FALSE, warning=FALSE}
#We will now read the file
mig_data <- read.csv("df.csv")

# create a theme for dot plots, which can be reused
theme_dotplot <- theme_bw(14) +
  theme(axis.text.y = element_text(size = rel(.75)),
        axis.ticks.y = element_blank(),
        axis.title.x = element_text(size = rel(.75)),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(size = 0.5),
        panel.grid.minor.x = element_blank())

colnames(mig_data)[9] <- "Gross_total_income"

# Combine 
mig_grouped <- as.data.table(mig_data)[, sum(Returns), by = .(Status, State_name)]

# assigning the second column name to a new name
colnames(mig_grouped)[3] <- "Total_returns"
colnames(mig_grouped)[2] <- "State_name"


# Filter plotting data
status_vec <- c("Inflow", "Outflow", "Same State")

mig_grouped$Total_returns <- as.numeric(mig_grouped$Total_returns) / 1000

mig_grouped <- mig_grouped %>% 
  filter(Status %in% status_vec) %>%
  filter(State_name != "Total US")
ggplot(mig_grouped,
       aes(Total_returns, fct_reorder2(State_name, Status, Total_returns, .desc = FALSE),
           color = Status)) +
  geom_point() + ggtitle("Total returns for top 10 states facet by year") + ylab("States") + xlab("Total Returns in thousands") +
  theme_dotplot +
  theme(axis.title=element_text(size=10),plot.title=element_text(size=13))


```



The following observations can be made from the above graph which is a facet graph of the number of tax returns of the migrant population with respect to the type of migration over the span of 5 years.

- The figures for states like Texas, California, New York, and Florida are high because their population size is significantly greater than those of other states.

- Texas is one of the very few states where the inflow migration is greater than the outflow migration. One of the many reasons could be low tax rates in Texas. Hence a majority of the migrants prefer to settle in Texas as compared to other states.

- Of all the states, the outflow migration is maximum in California. One of the many reasons for this trend could be the cost of living. It is far more expensive to live in California as compared to any other state in the United States. 



```{r, echo=FALSE, warning=FALSE}
#We will now read the file
mig_data <- read.csv("df.csv")

# create a theme for dot plots, which can be reused
theme_dotplot <- theme_bw(14) +
  theme(axis.text.y = element_text(size = rel(.75)),
        axis.ticks.y = element_blank(),
        axis.title.x = element_text(size = rel(.75)),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(size = 0.5),
        panel.grid.minor.x = element_blank())

colnames(mig_data)[9] <- "Gross_total_income"

# Combine 
mig_grouped <- as.data.table(mig_data)[, sum(Returns), by = .(Year, Status, State_name)]

# assigning the second column name to a new name
colnames(mig_grouped)[4] <- "Total_returns"
colnames(mig_grouped)[3] <- "State_name"

top_total_states <- c("California", "Texas", "Florida", "New York", "Georgia", "Virgina", "Ohio", "North Carolina", "Michigan", "Illinois")

# Filter plotting data
status_vec <- c("Inflow", "Outflow", "Same State")

mig_grouped <- mig_grouped %>% 
  filter(Status %in% status_vec) %>%
  filter(State_name != "Total US") %>%
  filter(State_name %in% top_total_states)

mig_grouped$Total_returns <- as.numeric(mig_grouped$Total_returns) / 1000

ggplot(mig_grouped,
       aes(Total_returns, fct_reorder2(State_name, Status, Total_returns, .desc = FALSE),
           color = Status)) +
  geom_point() + ggtitle("Total returns for top 10 states facet by year") + ylab("States") + xlab("Total Returns in thousands") +
  theme_dotplot +
  facet_wrap(~factor(Year)) +
  theme(axis.title=element_text(size=10),plot.title=element_text(size=13))

```





The above graph is a detailed analysis of the previous graph wherein we can see the different types of migration in the top 10 states.The following observations can be made from the above graph:

- In the top 10 states, the same state migration dominates the interstate migration.

- In states like Ohio and Michigan , the inflow and outflow migration are almost equal in numbers and this trend prevails over the span of 5 years from 2016 to 2020.




```{r, echo=FALSE, warning=FALSE}

```
